---
title: "Simple regressions NII on IR"
output:
  html_document:
    toc: yes
    toc_float: yes
  # pdf_document:
  #   toc: yes
---

```{=html}
<style>
  body {
  background-color: white;
  color: black;
}

pre, code {
  background-color: lightgray;
  color: grey;
}

pre:not([class]) {
  background-color: lightgray;
  color: black;
}

pre {
  border: 0px solid gray;
}

.hljs-string {
  color: darkgreen;
}

.hljs-number {
  color: blue;
}

.hljs-literal {
  color: brown;
}

.hljs-keyword {
  color: purple;
}

.tocify {
  border: 0px;
}

.list-group-item {
  background-color: lightgray;
  color: black;
}

.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
  background-color: blue;
  border-color: blue;
  color: white;
}

h1, h2, h3, h4, h5, h6 {
  color: black;
}

</style>
```
# Libraries

```{r}
library(readr)
library(dplyr)
library(stargazer)
library(forecast)
library(lubridate)
library(stats)
```

## Reading data

```{r}
data <- read_csv("data/for_regressing/interbank_nii.csv")

data_clean <- na.omit(data)
```

# Modeling

## First simple model

```{r}
model <- lm(NII ~ IR, data = data_clean)

stargazer(model, type = "text", align = TRUE, single.row = TRUE)
cat(paste("AIC: ", AIC(model)))
```

## First lagged model

```{r}
data_clean <- data_clean %>%
  mutate(IR_lag1 = lag(IR, 1),
         IR_lag2 = lag(IR, 2),
         IR_lag3 = lag(IR, 3))

model <- lm(NII ~ IR + IR_lag1 + IR_lag2 + IR_lag3, data = data_clean)

stargazer(model, type = "text", align = TRUE, single.row = TRUE)
cat(paste("AIC: ", AIC(model)))
```

This model is actually worse than the original one indicating that I need to somehow determine which exact lags to use.

# Correlation functions

From what I know usually people use these things to determine which lags to use, but I don't find them very useful. Still I decided to plot them anyways.

## ACF

```{r}
acf(data$IR, main="ACF of IR")
```

## PCF

```{r}
pacf(data$IR, main="PACF of IR")
```

# Best possible model

I dislike ACF and PCF and find them not very informative so instead I decided to iterate over all possible combinations of lags(up to 12) to find the best one.

## Best by AIC

```{r}
data$Date <- ymd(data$Date)

basic_model <- lm(NII ~ IR, data = data)

stargazer(basic_model, title = "Basic Model", type = "text", align = TRUE, single.row = TRUE)
cat(paste("AIC: ", AIC(basic_model)))

create_lags <- function(data, var, max_lag) {
  for (i in 1:max_lag) {
    data[paste0(var, "_lag", i)] <- lag(data[[var]], i)
  }
  return(data)
}

data_with_lags <- create_lags(data, "IR", 24)

test_lag_combinations <- function(data, max_lag) {
  best_model <- NULL
  best_aic <- Inf

  for (i in 1:max_lag) {
    for (combination in combn(1:max_lag, i, simplify = FALSE)) {
      formula <- as.formula(paste("NII ~", paste(paste0("IR_lag", combination), collapse = " + ")))
      model <- lm(formula, data = data)
      current_aic <- AIC(model)

      if (current_aic < best_aic) {
        best_model <- model
        best_aic <- current_aic
      }
    }
  }

  return(best_model)
}

best_model <- test_lag_combinations(data_with_lags, 24)

cat('\n\n')
stargazer(best_model, title = "Best model", type = "text", align = TRUE, single.row = TRUE)
cat(paste("AIC: ", AIC(best_model)))
cat('\n')
cat(paste("Best model formula:", deparse(formula(best_model))))
```

So the best model is the one with lag 1 and 3. Lag 1 makes sense as it is expected to have a big effect due to the delayed effect that such policy changes can have. Why Lag 3 is so good I do not understand.

```{r}
model <- lm(NII ~ IR + IR_lag1 + IR_lag3, data = data_with_lags)

stargazer(model, type = "text", align = TRUE, single.row = TRUE)
cat(paste("AIC: ", AIC(model)))
```

It's interesting that adding a non lagged IR actually results in a worse model if we compare AIC scores. This kind of makes sense as the effect of changes to IR is always delayed. But also if we compare the $R^2$ of different models then this model is better.

## Best by $R^2$

Just for fun I also checked what is the best model according to $R^2$. I know it's not designed for comparing models but still I thought we might see something interesting.

```{r}
data$Date <- ymd(data$Date)

# Basic regression
basic_model <- lm(NII ~ IR, data = data)

stargazer(basic_model, title = "Basic Model", type = "text", align = TRUE, single.row = TRUE)

# Function to create lagged variables
create_lags <- function(data, var, max_lag) {
  for (i in 1:max_lag) {
    data[paste0(var, "_lag", i)] <- lag(data[[var]], i)
  }
  return(data)
}

# Create lagged variables for IR
data_with_lags <- create_lags(data, "IR", 24)

# Function to test all possible lag combinations
test_lag_combinations <- function(data, max_lag) {
  best_model <- NULL
  best_r2 <- 0
  
  for (i in 1:max_lag) {
    for (combination in combn(1:max_lag, i, simplify = FALSE)) {
      formula <- as.formula(paste("NII ~", paste(paste0("IR_lag", combination), collapse = " + ")))
      model <- lm(formula, data = data)
      current_r2 <- summary(model)$r.squared
      
      if (current_r2 > best_r2) {
        best_model <- model
        best_r2 <- current_r2
      }
    }
  }
  
  return(best_model)
}

# Find the best model
best_model <- test_lag_combinations(data_with_lags, 12)

cat('\n\n')
stargazer(best_model, title = "Best model", type = "text", align = TRUE, single.row = TRUE)
cat('\n')
cat(paste("Best R-squared:", summary(best_model)$r.squared))
```

And we did see something interesting. Expectedly the best model is the one that has the most explanatory variables. But not quite. Even here the non-lagged IR is absent indicating that it is really not at all useful, meaning that there is indeed a very significant delay between policy changes taking effect and then affecting the actual results.



# HI

```{r}
data <- read_csv("data/relative/Interbank.csv")

data_adjusted <- data %>%
  arrange(Date) %>%
  mutate(IR_diff = IR - lag(IR, default = NA))

data_with_lags <- create_lags(data_adjusted, "IR", 24)
df <- data_with_lags[, c('Date', 'IR_lag7')]
# Save the modified dataframe to a CSV file
write.csv(df, "data/relative/Interbank_lag/Interbank_lag7.csv", row.names = FALSE)
```

