---
title: "Simple regressions NII on IR"
output:
  html_document:
    toc: yes
    toc_float: yes
  # pdf_document:
  #   toc: yes
---

```{=html}
<style>
  body {
  background-color: white;
  color: #d2d2d2;
}

pre, code {
  background-color: lightgray;
  color: black;
}

pre:not([class]) {
  background-color: lightgray;
  color: black;
}

pre {
  border: 0px solid gray;
}

.hljs-string {
  color: darkgreen;
}

.hljs-number {
  color: blue;
}

.hljs-literal {
  color: brown;
}

.hljs-keyword {
  color: purple;
}

.tocify {
  border: 0px;
}

.list-group-item {
  background-color: lightgray;
  color: black;
}

.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
  background-color: blue;
  border-color: blue;
  color: white;
}

h1, h2, h3, h4, h5, h6 {
  color: black;
}

</style>
```

------------------------------------------------------------------------

The [R plugin](https://www.jetbrains.com/help/pycharm/r-plugin-support.html) for IntelliJ-based IDEs provides handy capabilities to work with the [R Markdown](https://www.jetbrains.com/help/pycharm/r-markdown.html) files. To [add](https://www.jetbrains.com/help/pycharm/r-markdown.html#add-code-chunk) a new R chunk,

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
```

position the caret at any line or the code chunk, then click "+".

The code chunk appears:

```{r}
plot_median_quartiles <- function(file_path) {
  # Read the CSV file
  df <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Convert Date column to Date type
  df$Date <- as.Date(df$Date)
  
  # Ensure all columns are numeric or NA, excluding the Date column
  df_clean <- df %>%
    mutate(across(-Date, ~ as.numeric(as.character(.)), .names = "numeric_{col}"))
  
  # Drop non-numeric columns and keep only the Date and numeric columns
  df_numeric <- df_clean %>%
    select(Date, starts_with("numeric_")) %>%
    rename_with(~ sub("numeric_", "", .))
  
  # Convert to long format
  df_long <- df_numeric %>%
    pivot_longer(cols = -Date, names_to = "Bank", values_to = "Value")
  
  # Compute median and quartiles by date
  summary_stats <- df_long %>%
    group_by(Date) %>%
    summarise(
      Median = median(Value, na.rm = TRUE),
      Q1 = quantile(Value, 0.25, na.rm = TRUE),
      Q3 = quantile(Value, 0.75, na.rm = TRUE),
      .groups = 'drop'
    )
  
  # Plot using ggplot2
  ggplot(summary_stats, aes(x = Date)) +
    geom_line(aes(y = Median, color = "Median")) +
    geom_ribbon(aes(ymin = Q1, ymax = Q3), alpha = 0.2, fill = "grey20") +
    geom_vline(xintercept = as.Date("2018-01-01"), color = "red", linetype = "dashed") +
    geom_vline(xintercept = as.Date("2018-04-01"), color = "pink", linetype = "dashed") +
    labs(title = "Median and Quartiles of Values Over Time",
         x = "Date",
         y = "Value") +
    theme_minimal() +
    scale_color_manual(name = "Statistic", values = c("Median" = "blue")) +
    theme(legend.position = "top")
}
```

```{r}
plot_column <- function(filename, chosen_column) {
  # Read the CSV file
  data <- read.csv(filename, stringsAsFactors = FALSE)
  
  # Convert the Date column to Date type
  data$Date <- as.Date(data$Date)
  
  # Check if the chosen column exists
  if (!(chosen_column %in% colnames(data))) {
    stop("Chosen column ", chosen_column, " does not exist in the CSV file.")
  }
  
  # Plot the chosen column
  ggplot(data, aes(x = Date, y = get(chosen_column))) +
    geom_line() +
    geom_vline(xintercept = as.Date("2018-01-01"), color = "red", linetype = "dashed") +
    geom_vline(xintercept = as.Date("2018-04-01"), color = "pink", linetype = "dashed") +
    labs(title = paste("Time Series Plot of", chosen_column),
         x = "Date",
         y = chosen_column) +
    theme_minimal()
}
```

Type any R code in the chunk, for example:

```{r}
plot_column("../data/extracted/2018_to_now_monthly/refinanced_by_nbu.csv", "ОЩАДБАНК")
plot_column("../data/extracted/2018_to_now_monthly/securities.csv", "ОЩАДБАНК")
```

```{r}
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/relative/Interbank.csv', 'IR')

plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/relative/Policy.csv', 'IR')
```

```{r}
bank = "ПРИВАТБАНК"
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/2009_to_2017_quaterly/admin_expenses.csv', bank)
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/2009_to_2017_quaterly_diff/admin_expenses.csv', bank)
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/2018_to_now_monthly_differenced/admin_expenses.csv', bank)
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/admin_expenses.csv', bank)
```

```{r}
bank = "ОЩАДБАНК"
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/2009_to_2017_quaterly/admin_expenses.csv', bank)
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/2009_to_2017_quaterly_diff/admin_expenses.csv', bank)
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/2018_to_now_monthly_differenced/admin_expenses.csv', bank)
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/admin_expenses.csv', bank)
```

```{r}
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/admin_expenses.csv', "ПРИВАТБАНК")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/admin_expenses.csv', "ОЩАДБАНК")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/admin_expenses.csv', "УКРСИББАНК")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/admin_expenses.csv', "УКРЕКСІМБАНК")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/admin_expenses.csv', "УКРСОЦБАНК")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/admin_expenses.csv', "ПРОМІНВЕСТБАНК")
# plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/admin_expenses.csv', "НАДРА")
# plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/admin_expenses.csv', "ФОРУМ")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/admin_expenses.csv', "ПУМБ")
```

```{r}
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/net_commision_income.csv', "ПРИВАТБАНК")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/net_commision_income.csv', "ОЩАДБАНК")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/net_commision_income.csv', "УКРСИББАНК")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/net_commision_income.csv', "УКРЕКСІМБАНК")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/net_commision_income.csv', "УКРСОЦБАНК")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/net_commision_income.csv', "ПРОМІНВЕСТБАНК")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/net_commision_income.csv', "НАДРА")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/net_commision_income.csv', "ФОРУМ")
plot_column('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/net_commision_income.csv', "ПУМБ")
```

```{r}
plot_median_quartiles('~/Documents/GitHub/MonetaryPolicyEffectOnNetInterestMargins/data/extracted/complete/total_assets.csv')
```

```{r}
plot_median_quartiles("../data/relative/admin_expenses_to_total_assets.csv")
```

Now, click the **Run** button on the chunk toolbar to [execute](https://www.jetbrains.com/help/pycharm/r-markdown.html#run-r-code) the chunk code. The result should be placed under the chunk. Click the **Knit and Open Document** to build and preview an output.

```{r}
plot_median_quartiles("../data/relative/ovdp_to_total_assets.csv")
```

```{r}
plot_median_quartiles("../data/relative/net_commision_income_to_total_assets.csv")
```

```{r}
plot_median_quartiles("../data/relative/capital_to_total_assets.csv")
```

```{r}
plot_median_quartiles("../data/relative/net_interest_income_to_total_assets.csv")
```

```{r}
plot_median_quartiles("../data/relative/total_assets.csv")
```
